---
title: Regression Project
author: Connor Spencer, Lucas Koscho, Vishnu Jariwala
date: April 28th, 2025
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I pledge my honor that I have abided by the Stevens Honor System.

```{r}
# Set seed
CWID = 20030846
personal = CWID %% 10000
set.seed(personal)

# Load required packages
library(leaps)

# Load the data
insurance = read.csv('insurance.csv')

# Convert categories to factors
insurance$Region = as.factor(insurance$Region)
insurance$CarType = as.factor(insurance$CarType)
insurance$FuelType = as.factor(insurance$FuelType)
insurance$MaritalStatus = as.factor(insurance$MaritalStatus)

# Visualize data
summary(insurance)
```


```{r}
# Function for k-fold splitting
kfold = function(insuranceTest) {
  # Shuffle data
  insuranceTest = insuranceTest[sample(nrow(insuranceTest)), ]
  
  # Create 10 equally sized folds
  folds = cut(seq(1, nrow(insuranceTest)), breaks = 10, labels = FALSE)
  
  # Initialize list to store data for each fold
  fold_data = list()
  
  # Perform cross validation
  for (i in 1:10) {
    # Segment data by fold
    testIndexes = which(folds == i, arr.ind = TRUE)
    test_data = insuranceTest[testIndexes, ]
    train_data = insuranceTest[-testIndexes, ]
    
    # Store train and test data for each fold
    fold_data[[i]] = list(train_data = train_data, test_data = test_data)
  }
  return(fold_data)
}
```

```{r}
# Use k-fold splitting on insurance data
split_data = kfold(insurance)

# View training and test data for first fold
summary(split_data[[1]]$train_data)
summary(split_data[[1]]$test_data)
```

```{r}
# Store all full models
full_models = list()

# Initialize R-squared for full model
R_squared_full = numeric(length(split_data))

# Initialize AIC for full model
AIC_full = numeric(length(split_data))

# Initialize RMSE for full model
rmse_full = numeric(length(split_data))

# Access training and test data for each fold
for (i in 1:length(split_data)) {
  train_data = split_data[[i]]$train_data
  test_data = split_data[[i]]$test_data
  
  # Fit full model using all predictors
  full_model = lm(AnnualPremium ~ ., data = train_data)
  full_models[[i]] = full_model
  R_squared_full[i] = summary(full_model)$adj.r.squared
  AIC_full[i] = AIC(full_model)
  
  # Test full model against test data
  pred_test_full = predict(full_model, newdata = test_data)
  rmse_full[i] = sqrt(mean((pred_test_full-test_data$AnnualPremium)^2))
}
summary(full_models[[1]])
```

```{r}
# Store all stepwise models
step_models = list()

# Initialize R-squared for stepwise model
R_squared_step = numeric(length(split_data))

# Initialize AIC for stepwise model
AIC_step = numeric(length(split_data))

# Initialize RMSE for stepwise model
rmse_step = numeric(length(split_data))

# Access training and test data for each fold
for (i in 1:length(split_data)) {
  train_data = split_data[[i]]$train_data
  test_data = split_data[[i]]$test_data
  
  # Fit full model using all predictors
  full_model = lm(AnnualPremium ~ ., data = train_data)
  
   # Stepwise selection
  step_model = step(full_model, direction = 'both', trace = FALSE)
  step_models[[i]] = step_model
  R_squared_step[i] = summary(step_model)$adj.r.squared
  AIC_step[i] = AIC(step_model)
  
  # Test stepwise model against test data
  pred_test_step = predict(step_model, newdata = test_data)
  rmse_step[i] = sqrt(mean((pred_test_step-test_data$AnnualPremium)^2))
}
summary(step_models[[1]])
```

```{r}
# Store all selected models
models_selected = list()

# Initialize R-squared for selected model
R_squared_select = numeric(length(split_data))

# Initialize AIC for selected model
AIC_select = numeric(length(split_data))

# Initialize RMSE for selected model
rmse_select = numeric(length(split_data))

# Access training and test data for each fold
for (i in 1:length(split_data)) {
  train_data = split_data[[i]]$train_data
  test_data = split_data[[i]]$test_data
  
  # Regsubsets to find best subset of predictors
  subset_model = regsubsets(AnnualPremium ~ ., data = train_data, nvmax = ncol(train_data)-1)
  reg_summary = summary(subset_model)
  num_predictors = which.max(reg_summary$adjr2)
  
  # Find which predictors were selected
  selected_predict = names(which(reg_summary$which[num_predictors, -1]))
  selected_predict = selected_predict[selected_predict %in% names(train_data)]
  
  # Refit model using selected predictors
  formula_selected = as.formula(paste('AnnualPremium ~', paste(selected_predict, collapse = ' + ')))
  model_selected = lm(formula_selected, data = train_data)
  models_selected[[i]] = model_selected
  R_squared_select[i] = summary(model_selected)$adj.r.squared
  AIC_select[i] = AIC(model_selected)

  # Test selected model against test data
  pred_test_select = predict(model_selected, newdata = test_data)
  rmse_select[i] = sqrt(mean((pred_test_select-test_data$AnnualPremium)^2))
}
summary(models_selected[[1]])
```

```{r}
# Compare R-squared across models
mean(R_squared_full)
mean(R_squared_step)
mean(R_squared_select)
print('')

# Compare AIC across models
mean(AIC_full)
mean(AIC_step)
mean(AIC_select)
print('')

# Compare RMSE across models
mean(rmse_full)
mean(rmse_step)
mean(rmse_select)
```

### The Stepwise Selection Model has the highest R-squared value (0.8291), the lowest AIC (3974.43), and the second lowest root Mean Squared Error (136.27). Therefore, we will be using this as our model.

```{r}
# Residual Plot
plot(
  step_models[[1]]$fitted.values,
  step_models[[1]]$residuals,
  xlab = "Fitted Values",
  ylab = "Residuals",
  main = "Residuals vs Fitted Values",
  pch = 20,
  col = "blue"
)
abline(h = 0, lty = 2, col = "red")
```

Since our residuals on the graph create no pattern, it tells us that our linearity and variance of our model is accurate.

```{r}
# QQ Plot
qqnorm(step_models[[1]]$residuals, pch = 20, main = "Normal Q-Q Plot")
qqline(step_models[[1]]$residuals, col = "red", lwd = 2)
```

The plot follows a mostly straight line, with only slight deviations at the tails, indicating that there is normality among our residuals.